Capacity and Utilization analysis of Toronto shelter(2024)

### 2024-09-27 paper/paper_Li.qmd
Add files via upload

With respect to the structure of this paper, @sec-data includes an overview of the data and @sec-model show the results of model; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.

@@

The dataset used in this analysis is from The "Daily Shelter & Overnight Service Usage" page , this website offers daily data of the occupancy and capacity of Toronto’s shelter system, including how many shelter beds are being used and the total available capacity, updated everyday except on weekends and holidays. The variables or measurements included in this analysis are "Bed Based Capacity", which refer to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" refer to ccupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.

Another dataset called "toronto-shelter-system-flow" also includes data about Toronto shelters, but it is more focused on the characteristics of people living in shelters, not the macro data so I did not choose it.

The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@thereferencecanbewhatever] package were used to simulate the dataset and generate tests for it. The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@thereferencecanbewhatever] packages were then applied in order to download the raw Toronto Public shelter dataset. Next, the `tidyverse` package [@thereferencecanbewhatever] was used to clean the raw dataset and test the cleaned dataset.

@@

@tbl-sample show a sample of cleaned occupancy data, Since the statistical record formats of different shelters are different, each row contains different information marked as NA, but this does not negattive affect the analysis.
\newpage

# Result {#sec-model}

@@

# Plot the occupancy rate trend
ggplot(daily_occupancy_rate, aes(x = OCCUPANCY_DATE)) +
  geom_line(aes(y = Bed_Occupancy_Rate, color = "Bed Occupancy Rate"), size = 1) +
  geom_line(aes(y = Room_Occupancy_Rate, color = "Room Occupancy Rate"), size = 1, linetype = "dashed") +
  labs(title = "", x = "Date", y = "Occupancy Rate (%)") +
  theme_minimal() +
  scale_color_manual(values = c("Bed Occupancy Rate" = "blue", "Room Occupancy Rate" = "red")) +
  theme(legend.title = element_blank())
```
@fig-occupancy-rates shows the occupancy rates for beds and rooms in the shelter system over time. The blue line represents the bed occupancy rate, while the red line represents the room occupancy rate. Overall, the room occupancy rate (red) is higher and more stable, staying close to 100%. In contrast, the bed occupancy rate (blue) shows more fluctuations, generally ranging between 98% and 99%.

@@

@fig-occupancy-capital-relat displays the average occupancy rate across various shelters, categorized by sector: Men, Women, Mixed Adult, and Youth. The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters. Each horizontal blue bar corresponds to the average occupancy rate of a specific shelter. Most shelters have an occupancy rate close to 100%, indicating that their beds are almost fully or completely occupied. A few shelters have occupancy rates below 75%, showing that they have relatively lower bed utilization.

>>>>

Concerning the structure of this paper, @sec-data includes an overview and visualization of the data; @sec-discussion includes a discussion of the results; and @sec-appendix includes supplementary information.

@@

The dataset used in this analysis is from The "Daily Shelter & Overnight Service Usage" page, this website offers daily data on the occupancy and capacity of Toronto’s shelter system, including how many shelter beds are being used and the total available capacity, updated every day except on weekends and holidays. The variables or measurements included in this analysis are "Bed Based Capacity", which refers to Occupancy is measured by the number of beds, typically for programs with shared sleeping areas; "Room Based Capacity" to occupancy is measured by the number of rooms, usually for family programs or hotel-based programs where rooms are not shared between households.

Another dataset called "toronto-shelter-system-flow" also includes data about Toronto shelters. Still, it is more focused on the characteristics of people living in shelters, not the macro data so I did not choose it.

The project is structure base on Rohan's principes [@rohan], and used the **R** programming language [@citeR], the `tidyverse` [@citetidyverse] package were used to simulate the dataset and generate tests for it. The `opendatatoronto` [@citeopendatatoronto] and `tidyverse` [@citetidyverse] packages were then applied in order to download the raw Toronto Public shelter dataset. Next, the `tidyverse` package [@citetidyverse] was used to clean the raw dataset and test the cleaned dataset. The `knitr` packages help illustrate the sample data (Xie 2024)

@@

@tbl-sample show a sample of cleaned occupancy data, Since the statistical record formats of different shelters are different, each row contains different information marked as NA, but this does not negatively affect the analysis.
\newpage

## analysis {#sec-data-analysis}

@@

# Plot the occupancy rate trend
ggplot(daily_occupancy_rate, aes(x = OCCUPANCY_DATE)) +
  geom_line(aes(y = Bed_Occupancy_Rate, color = "Bed Occupancy Rate"), size = 1) +
  geom_line(aes(y = Room_Occupancy_Rate, color = "Room Occupancy Rate"), size = 1, linetype = "dashed") +
  labs(title = "", x = "Date", y = "Occupancy Rate (%)") +
  theme_minimal() +
  scale_color_manual(values = c("Bed Occupancy Rate" = "blue", "Room Occupancy Rate" = "red")) +
  theme(legend.position = "top", legend.title = element_blank())
```
@fig-occupancy-rates shows the occupancy rates for beds and rooms in the shelter system over time. The blue line represents the bed occupancy rate, while the red line represents the room occupancy rate. Overall, the room occupancy rate (red) is higher and more stable, staying close to 100%. In contrast, the bed occupancy rate (blue) shows more fluctuations, generally between 98% and 99%.

@@

@fig-occupancy-capital-relat displays the average occupancy rate across various shelters, categorized by sector: Men, Women, Mixed Adults, and Youth. The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters. Each horizontal blue bar corresponds to the average occupancy rate of a specific shelter. Most shelters have an occupancy rate close to 100%, indicating that their beds are almost fully or completely occupied. A few shelters have occupancy rates below 75%, showing that they have relatively lower bed utilization.
